|

### 2023-02-01 paper.Rmd
Wed Feb  1 14:00:06 CET 2023: update-remote.sh

```{python read_runtime_vars, echo=FALSE}
from json import load
D_AIRFOIL_2hl = load(
  open("/Users/rmn/github/master_thesis/data/airfoil_2hl_150hn_fullRun_30gens/summary.json", "r", encoding="utf-8")
  )

@@

```{r runtime_vars, echo=FALSE}

@@

@@

The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE of `r round(, 3)` (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness. Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.

@@

The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure  \@ref(fig:airfoil-distribution-fitness-30gens). The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP on both sets since it shows a lower median RMSE as well as less dispersion of the solutions measured by the standard deviation. An interesting observation from studying the distribution of final fitness scores for both algorithms is that while both algorithms have a very similiar interquartile range (similiar box shape), pre-trained DAE-GP has more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \@ref(fig:airfoil-median-fitness-30gens).

Since previous work by @dae-gp_2022_symreg demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP. Figure \@ref(fig:airfoil-size-combined-30gens) shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population. It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search. Regarding the size of the best solution per generation, the results show that the pre-trained version of algorithm did produce solutions with the same median size as traditional DAE-GP.

@@

```{r runtime_vars, echo=FALSE}

@@

Suprisingly, even after correcting the runtime for both DAE-GP results by removing the time spend during sampling, my implementation of pre-training for DAE-GP still on average takes more time to finish the search process as shown in figure \@ref(fig:airfoil-runtime-noSampling-total). While traditional DAE-GP has a median runtime without sampling time of `r format(REG_TIME_NOSAMLING, scientific=F)` seconds , the runtime of pre-trained DAE-GP, without time for samling and pre-training, still increases by a factor of `r round(PT_TIME_NOSAMPLING_NOPT / REG_TIME_NOSAMLING, 3)` to `r format(PT_TIME_NOSAMPLING_NOPT, scientific=F)` seconds. While investigating other reasons for the differences in runtime, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models ( `r format(REG_PRAMS, scientific=F)`).

>>>>

```{python read_solQual_vars, echo=FALSE}
from json import load
D_AIRFOIL_2hl = load(
  open("/Users/rmn/github/master_thesis/data/airfoil_2hl_150hn_fullRun_30gens_withGP/summary.json", "r", encoding="utf-8")
  )

@@

```{r solQual_vars, echo=FALSE}

@@

REG_MED_BESTSIZE <- py$D_AIRFOIL_2hl$REG_MED_BESTSIZE
PT_MED_BESTSIZE <- py$D_AIRFOIL_2hl$PT_MED_BESTSIZE
GP_MED_BESTSIZE <- py$D_AIRFOIL_2hl$GP_MED_BESTSIZE

@@

The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness. Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.

@@

The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure  \@ref(fig:airfoil-distribution-fitness-30gens). The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP  since it shows a lower median RMSE of `r round(REG_TEST_MED_FF, 3)` for the test set (`r round(PT_TEST_MED_FF, 3)` for pre-trained DAE-GP) as well as less dispersion of the solutions measured by a standard deviation of `r round(REG_TEST_STDD_FF, 3)` (`r round(PT_TEST_STDD_FF, 3)` for pre-trained DAE-GP). For reference, the median final RMSE for the test set of standard GP is at `r round(GP_TEST_MED_FF, 3)` which is larger by a factor of `r round(GP_TEST_MED_FF / PT_TEST_MED_FF, 3)` if compared to pre-trained DAE-GP.

An interesting observation from studying the distribution of final fitness scores for both algorithms is that while both algorithms have a very similiar interquartile range (similar box shape), pre-trained DAE-GP has more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \@ref(fig:airfoil-median-fitness-30gens).

Since previous work by @dae-gp_2022_symreg demonstrated that DAE-GP, for a given number of 10.000 fitness evaluations, is able to produce solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP. Figure \@ref(fig:airfoil-size-combined-30gens) shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population. It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search. Regarding the size of the best solution per generation, the results show that the pre-trained version of algorithm did produce solutions with the same median size as traditional DAE-GP. The median size

@@

```{r runtime_vars, include=FALSE}

@@

Suprisingly, even after correcting the runtime for both DAE-GP results by removing the time spend during sampling, my implementation of pre-training for DAE-GP still on average takes more time to finish the search process as shown in figure \@ref(fig:airfoil-runtime-noSampling-total). While traditional DAE-GP has a median runtime without sampling time of `r format(REG_TIME_NOSAMLING, scientific=F)` seconds , the runtime of pre-trained DAE-GP, without time for samling and pre-training, still increases by a factor of `r round(PT_TIME_NOSAMPLING_NOPT / REG_TIME_NOSAMLING, 3)` to `r format(PT_TIME_NOSAMPLING_NOPT, scientific=F)` seconds. While investigating other reasons for the differences in runtime, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models ( `r format(REG_PARAMS, scientific=F)`).

### 2023-02-01 paper.tex
Wed Feb  1 14:00:06 CET 2023: update-remote.sh

Johannes Gutenberg University Mainz}
\author{}
\date{\vspace{-2.5em}Date of Submission: 2023-01-31}

@@

The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness.Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.

@@

The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure \ref{fig:airfoil-distribution-fitness-30gens}. The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP on both sets since it shows a lower median RMSE as well as less dispersion of the solutions measured by the standard deviation. An interesting observation from studying the distribution of final fitness scores for both algorithms is that while both algorithms have a very similiar interquartile range (similiar box shape), pre-trained DAE-GP has more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \ref{fig:airfoil-median-fitness-30gens}.

@@

Suprinsingly, as shown in figure \ref{fig:airfoil-runtime-total}, for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median runtime of DAE-GP. If the time spend on pre-training is included into the total runtime, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of 2.897 from a median runtime of 5076.448 seconds for regular DAE-GP to 14706.47 seconds for the pre-trained algorithm. Even in the case where the time spend for pre-training is excluded from the total runtime, pre-training still increases the runtime of DAE-GP by a factor of 2.47 to 12536.89 seconds.

@@

Suprisingly, even after correcting the runtime for both DAE-GP results by removing the time spend during sampling, my implementation of pre-training for DAE-GP still on average takes more time to finish the search process as shown in figure \ref{fig:airfoil-runtime-noSampling-total}. While traditional DAE-GP has a median runtime without sampling time of 3550.277 seconds , the runtime of pre-trained DAE-GP, without time for samling and pre-training, still increases by a factor of 1.865 to 6622.027 seconds.

>>>>

Johannes Gutenberg University Mainz}
\author{}
\date{\vspace{-2.5em}Date of Submission: 2023-02-01}

@@

The first interesting observation from this experiment is, that the pre-trained DAE-GP shows slightly lower mean RMSE of (higher fitness) but a higher median RMSE (lower fitness) than regular DAE-GP for both the training and the testing set in overall fitness. Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behavior.

@@

The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure \ref{fig:airfoil-distribution-fitness-30gens}. The distribution for regular DAE-GP shows to be preferable to that of pre-trained DAE-GP since it shows a lower median RMSE of 6.991 for the test set (7.68 for pre-trained DAE-GP) as well as less dispersion of the solutions measured by a standard deviation of 11.989 (12.548 for pre-trained DAE-GP). An interesting observation from studying the distribution of final fitness scores for both algorithms is that while both algorithms have a very similiar interquartile range (similiar box shape), pre-trained DAE-GP has more extreme outliers for high RMSE which explains the slightly worse performance if looking at median best fitness present in figure \ref{fig:airfoil-median-fitness-30gens}.

@@

Suprinsingly, as shown in figure \ref{fig:airfoil-runtime-total}, for the experiment conducted on the airfoil dataset with 2 hidden layers a strong negative impact of pre-training can be observed for the median runtime of DAE-GP. If the time spend on pre-training is included into the total runtime, the median duration for pre-trained DAE-GP over all 10 runs increases by a factor of 3.121 from a median runtime of 4712.303 seconds for regular DAE-GP to 14706.47 seconds for the pre-trained algorithm. Even in the case where the time spend for pre-training is excluded from the total runtime, pre-training still increases the runtime of DAE-GP by a factor of 2.66 to 12536.89 seconds.

@@

Suprisingly, even after correcting the runtime for both DAE-GP results by removing the time spend during sampling, my implementation of pre-training for DAE-GP still on average takes more time to finish the search process as shown in figure \ref{fig:airfoil-runtime-noSampling-total}. While traditional DAE-GP has a median runtime without sampling time of 3550.277 seconds , the runtime of pre-trained DAE-GP, without time for samling and pre-training, still increases by a factor of 1.865 to 6622.027 seconds. While investigating other reasons for the differences in runtime, I verified that both DAE-GP variants use the exact same number of trainable parameters for each generations DAE-LSTMS models ( NULL).
